
\newglossaryentry{service}
{
name={Service},
text={service},
description={A service consists in a set of functionalities, typically
using one or a few different resources, which allows a complete
business goal. For example, a billing service, a user management
service, in a shopping application}
}

\newglossaryentry{DevOps}
{
name={DevOps},
description={DevOps is a
portmanteau word composed of (software) development and (IT)
operations. DevOps are mainly a set of practices that combines those
in order to produce faster better products for customers. Globally, it
enhances the application lifecycle through methods used during the
lifecycle phases, plan, develop, deliver and operate.}
}

\newglossaryentry{p2p} {
name={Peer-to-peer},
text={peer-to-peer},
description={Peer-to-peer
computing is a type of architecture for distributed applications to
spread their workloads between peers of the system on which it is
deployed upon. Peers all have the same privileges and same function
abilities.}
}

\newglossaryentry{tls} {
name={Transport Layer Security},
description={TLS is a security protocol that provides secured and private
communications in computer networks. It used for HTTPS, but also emails,
instant messaging, etc.}
}


\newglossaryentry{api} {
name={Application Programming Interface},
description={An API is a software interface allowing other programs to
interact with the software functionalities that are exposed by this interface.
One of the goals is to keep the software internals hidden and expose only what
is relevant for software communications.}
}

\newglossaryentry{server} {
name={Server},
text={server},
description={A server is a \emph{computer} designed to process requests,
offer functionalities and deliver data to other, possibly multiple softwares or
devices that are called clients.}  }

\newglossaryentry{dc} {
name={Data center},
text={data center},
description={A data center
is a physical facility hosting networked computing and storage
components.}
}

\newglossaryentry{udp} {
name={User Datagram Protocol},
description={UDP is one of the communication protocol over IP network. Contrary to TCP, UDP does not require to establish a connection between the client and the server.}
}

\newglossaryentry{tcp} {
name={Transmission Control Protocol},
description={TCP is one of the main communication protocol,
complemented by the IP (Internet Protocol). TCP is based on the
establishment of a connection between the client and the server before
data are sent to deliver reliable data streams.}
}

\newglossaryentry{iot} {
name={Internet of things},
description={The internet of things consists in physical devices that carry sensors, have processing capabilities, and different technologies, to connect to other devices and computing systems, not necessarily on the internet. For examples, smart homes can be composed of different devices that compose the internet of things, like cameras, thermostats, etc.}
}


\newglossaryentry{router} {
name={Router},
text={router},
description={A router is a network device dedicated to forwarding data packets between networks. On the internet, they direct the traffic where it is needed.}
}

\newglossaryentry{switch} {
name={Switch},
text={switch},
description={A switch is a network device dedicated to forwarding data between the devices connected to it.}
}

% HERE

\newglossaryentry{disk array} {
name={Disk array},
text={disk array},
description={A disk array is a set of disk drives that compose a disk storage system.}
}

\newglossaryentry{loosely coupled} {
name={Loosely coupled},
text={loosely coupled},
description={Loose coupling refers to a model where the components of the system are connected but not heavily dependent, so they are not affected by changes or failures of others.}
}

\newglossaryentry{Contentdn} {
name={Content Delivery Network},
description={CDNs are a layer in the internet ecosystem. They consist in geo-distributed networks of data centers that delivers content (such as images, softwares, videos, streaming media, etc.) to end-users by being close to them. For example, Netflix or Youtube use CDN in most countries to deliver their videos faster.}
}

\newglossaryentry{cache} {
name={Cache},
text={cache},
description={A cache is a component that stores data closer to its usage point so future requests on this data will be served faster. For example, your browser saves logos or images of website so it does not need to download them again next time you are navigating them.}
}

\newglossaryentry{Domainns} {
name={Domain Name System},
description={The DNS is a naming system used to identify computers by a human friendly domain name by mapping it to IP addresses. For example, 8.8.8.8 is the primary DNS server for Google DNS.}
}

\newglossaryentry{cloudlet} {
name={Cloudlet},
text={cloudlet},
description={A cloudlet is a small-scale data center that is mobile at the edge of the internet.}
}

\newglossaryentry{smart contract} {
name={Smart contract},
text={smart contract},
description={A smart contract is a class of computer protocol that allows, check and execute the actions required by the terms of a contract. The main idea is to have automated agreements to ensure the terms of smart contract terms depending of predetermined factors.}
}

\newglossaryentry{hypervisor} {
name={Hypervisor},
text={hypervisor},
description={A hypervisor is a software that can create and run virtual machines (virtualization of computer systems).}
}

\newglossaryentry{microservice} {
name={Microservice},
text={microservice},
description={The microservice architecture is an architectural style that composes an application as a collection of services thatare loosely coupled, deployable independently, and execute each a small and coherent part of the business functions.}
}
