\chapter*{Acknowledgment}

There are a lot of people to thank for my work in my thesis.
%
A lot of really great persons, who taught me a lot, whether it was
directly related to my work, or for transversal knowledge.
%
At some point, I might not be able to describe how useful they were
and how lucky I was to talk with them, in good enough terms, but I
will try anyway.

First of all, I would like to thank my
supervisors.
%
Adrien Lebre, my ``official'' supervisor, was the one to pull me into
this great adventure that is a PhD thesis.
%
I saw firsthand how difficult it can be, and I would never have gotten
into it with anyone else.
%
I knew how dedicated he is with his other PhD students, and though it
was not always easy, of course, he helped and supported me however he
could, and pushed me whenever needed to give my best in my work.
%
Ronan-Alexandre Cherrueau, who definitely was my supervisor even
though it never was official, for the first year and a half.
%
Thanks for your really good ideas, pedagogy, and support.

I would also like to thank everyone in my defense jury, Sara
Bouchenak, Noël De Palma, Thomas Ledoux and Pierre Sens, first for
their time and second for their great questions and remarks during my
defense.

Then, all the people who participated in the project at some point.
%
Matthieu Simonin and Javier Rojas Balderrama, for Openstackoïd and all
the huge help in the beginning of my thesis.
%
David Espinel Sarmiento, with his thesis which initiated the Cross
collaboration, and great input on how it works, and a great partner in
sea kayak.
%
Karim Manaouil and Rodolfo, who participated in the Cheops meetings we
had in spring 2021.
%
Especially thanks to Karim with his help on Kubernetes and Cross and
always had a great perspective on different things.
%
My interns from spring 2021, Matthieu Juzdzewski and Arnaud Szymanek,
who were a great help to understand Consul and how Cheops should be
implemented, and were really great to supervise.
%
Last, but definitely not least, Geo Johns Antony, for all his great
knowledge and input, his patience, and the draft of the collaboration
model.
%
I wish him all the best for the end of his own thesis, even though I
am going to be around to try and help.


Then comes the people from the STACK team.
%
I will try to avoid thanking everyone by name, but some stand out.
%
Anthony Simonet, who took me as his intern one day, and was a great
supervisor and a huge help.
%
Johnathan Pastor, for his work on ROME, without which I would not have
been in the team either, who was always a good friend and a really
nice person, with insightful knowledge that helped a lot during all
these years.
%
Thuy Linh Nguyen and Jad Darrous, who were not only friends, but
convinced me someday to pursue a PhD thesis.
%
Hélène Coullon who was always a good friend too, and one of the nicest
person I had the chance to meet.
%
Maverick Chardet, Émile Cadorel, Fatima-Zahra Boujdad also need to be
thanked, and more recently, Antoine Omond and Baptiste Jonglez.


I am not going to mention every single one of my teachers, but I would
like to thank them anyway.
%
In particular, I would like to thank my university teachers from my
first Licence in Université du Havre Fabrice Durand and Véronique Jay
and of course all of my teachers in my computer science degrees from
Université de Nantes, with a special mention to Florian Richoux,
Frédéric Goualard and Hala Skaf-Molli who really inspired me during
the long road to PhD student.

Because a thesis would never happen without people from the
administrative part, I would like to thank everyone who helped me to
achieve mine.
%
Thanks to Anne-Claire Binetruy and Catherine Fourny to help me with
Inria and IMT Atlantique relations and being extra patient with my
inability to deal with administration.
%
Thanks to Michelle Dauvé for her work during my first year, and more
to Delphine Turlier, who was so helpful and always available for my
strange questions.

I need also to ``big-up'' my friends, from Université de
Nantes, Charlène, Charles-Éric and Sylvain, but also Samuel, Dorian,
Robin.
%
Nathalie, from Université du Havre, who supported me those long years,
from afar.
%
The Garithos crew, especially Logan and Brandon, who took all my
complaints these three years without complaining themselves.
%
The PhD students discord (as the people in it, obviously) also has its
own place here, both for moral support and help to solve problems, and
coincidentally celebrated its 3rd birthday on the day of my
defense. Happy Birthday, may you forever be a beacon for PhD students.


Finally, I am going to fall even more in the clichés by thanking my
family for their everlasting support throughout the years.
%
In particular, I would like to say thank you to my love Florent, who
is always my greatest source of support, inspiration and knowledge,
and has been for almost twenty years.
